---
title       : Pretty Graphs in R
subtitle    : A Realistic Approach
author      : Steve Bronder
job         : Marketing Analyst 
framework   : revealjs        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : github      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
ext_widgets: {rCharts: "libraries/highcharts"}
revealjs:
  theme: Simple
--- 

### Pretty Graphs in R


 A Realistic Approach
 \- Steve Bronder

--- 

### Outline

<ul class='incremental'>
  <li class='fragment'>What is R?</li>
  <li class='fragment'>Why and why not R?</li>
  <li class='fragment'>Describing the Language</li>
  <li class='fragment'>Data manipulation examples</li>
  <li class='fragment'>Pretty graph examples</li>
  <li class='fragment'>NLP in R</li>
  <li class='fragment'>Learning Resources</li>
</ul>

---

### What is R?

<ul class='incremental'>
  <li class='fragment'>Statistical language for data analysis</li>
    <li class='fragment'>Simple and effective language </li>
  <li class='fragment'>Can talk to Python, Julia, C, C++, HTML, CSS, etc.</li>
  <li class='fragment'>Extended by user contributions</li>
</ul>


---  .class &vertical 

### Why useR?



<ul class='incremental'>
  <li class='fragment'>Created by Statisticians</li>
  <li class='fragment'>Strong Community</li>
  <li class='fragment'>Being lazy. work smart and efficient</li>
  <li class='fragment'>Easy to prototype and make actionable</li>
  <li class='fragment'>Nothing beats R in Graphics</li>
</ul>


<script>
$('ul.incremental li').addClass('fragment')
$('ol.incremental li').addClass('fragment')
</script>

***

### Why notR?

<ul class='incremental'>
  <li class='fragment'>Created by Statisticians</li>
  <li class='fragment'>Sharp learning curve</li>
  <li class='fragment'>No more point and clicking</li>
  <li class='fragment'>Funky syntax</li>
   <li class='fragment'>Poorly written R code is terribly slow</li>
</ul>


<script>
$('ul.incremental li').addClass('fragment')
$('ol.incremental li').addClass('fragment')
</script>

***

### Fast and Slow R Code

```{r fastnSlow,cache=FALSE}

library(microbenchmark)

x <- runif(100)
microbenchmark(
  sqrt(x),
  x ^ 0.5
)
```


--- .class &vertical 

### Describing the Language
##### Everything is a function

```{r,cache=FALSE, class="fragment"}
1 + 1
`+`(1,1)
```


***

### Describing the Language
##### Make Statistics Easy

```{r,cache=FALSE,results='asis'}
data(iris)
fit.1 <- lm(Sepal.Length ~ Petal.Length + Species, data = iris)
```

```{r,cache=FALSE,results='asis',echo=FALSE}
stargazer::stargazer(fit.1,type = "html",
          single.row = TRUE,font.size="small",title = "Regression Results",notes="")
```

---  .class &vertical 

### Describing the Language
##### informal and formal OO systems available

```{r}
library(pryr)

# make a generic class and throw random things into it
foo <- structure(list(), class = "foo")
foo$balance <- 100
foo$withdraw <- function(x){
  foo$balance <<- foo$balance - x
}
foo$deposit <- function(x){
  foo$balance <<- foo$balance + x
}

foo$withdraw(10)
foo$deposit(1000)
foo$balance

otype(foo)

```


*** 

### Describing the language
##### informal and formal OO systems available

```{r refstuff,cache=FALSE}
library(methods)
# make a formal RC class
Accounts <- setRefClass("Accounts",
  fields = list(balance = "numeric"),
  methods = list(
    withdraw = function(x) {
      balance <<- balance - x
    },
    deposit = function(x) {
      balance <<- balance + x
    }
  )
)

a <- Accounts$new(balance = 100)
a$deposit(100)
a$balance
a$lottery <- "Insert winning lottery numbers here"
otype(a)

```


--- .class &vertical 

### Data manipulation in R
##### data.table Objective

<ol class='incremental'>
  <li class='fragment'>Develop light, SQL style data manipulation language in R.</li>
    <li class='fragment'>Fast, readable, consistent </li>
</ol>


***

### Data manipulation in R

- Create bins by flower type for sepal length

```{r, cache=FALSE, class = 'fragment'}
library(reshape2)
data(iris)
head(iris)
iris$cut.sepal <- melt(vapply(split(iris$Sepal.Length,iris$Species),
                   function(x) cut(x,5),vector(mode = "integer", length=50)))$value
head(iris,n=5)
```

***

### Data manipulation in R
##### data.table Answer

  `data[i , j, by]`

| Syntax  | SQL     | Meaning             | Example                         |
| ------- |:-------:|:-------------------:| -------------------------------:|
| i       |WHERE    | subset rows by i    | price > 0                       |
| j       |SELECT   | using column j      | logPrice := log(price)          |
| by      |GROUP-BY | grouped by          | by = store                      |

```{r,cache=FALSE, echo=FALSE}
iris$cut.sepal <- NULL
```

```{r,cache=FALSE}
iris <- data.table(iris)
iris[Petal.Width >=.2 , cut.sepal := as.integer(cut(Sepal.Length, 5)), by = Species]
```

--- .class &vertical

### Example: Cleaning Data
General Payment Dataset:
> - All general payments made to physicians and hospitals 
> - Recipients name, address, and specialty, payment, and "nature of payment", etc.
> - 2.6 million by 63 variables

```{r dataTableExample, cache = FALSE,cache.lazy = FALSE,warning=FALSE}
library(data.table, quietly =  TRUE,warn.conflicts = FALSE)
dat <- fread("/home/steve/Documents/OP_DTL_GNRL_PGYR2013_P06302015.csv")
```
***

### Example: Aggregation

```{r dataTableExample2Fake, cache=FALSE,cache.lazy = FALSE,warning=FALSE,eval=FALSE}
#Remove $ sign from Dollars column
  dat[ , Total_Amount_of_Payment_USDollars := as.numeric(substring(Total_Amount_of_Payment_USDollars,2))]

# Set the key variable for taking unique values
setkey(dat, Recipient_State)

#Get the averages for each state
  dat.means <- unique(dat[, mean(Total_Amount_of_Payment_USDollars,
                                           na.rm = TRUE),by = Recipient_State])
```

<!---
This actually runs
-->
```{r dataTableExample2, cache=FALSE,cache.lazy = FALSE,warning=FALSE}
#Remove $ sign from Dollars column
invisible(
  dat[ , Total_Amount_of_Payment_USDollars := as.numeric(substring(Total_Amount_of_Payment_USDollars,2))]
  )

# Set the key variable for taking unique values
setkey(dat, Recipient_State)

#Get the averages for each state
system.time(
  dat.means <- unique(dat[, mean(Total_Amount_of_Payment_USDollars,
                                           na.rm = TRUE),by = Recipient_State])
  )
rm(dat)
```



***

### Example: Displaying Data

```{r hiddendataTableEx, cache=FALSE,eval=FALSE,echo=FALSE,warning=FALSE,message=FALSE}
setnames(dat.means,"V1","payments")
colnames(dat.means) <- c("state","payments")
```

```{r dataTableExample4, cache = FALSE,warning=FALSE,echo=FALSE,comment=NA,results='asis'}
dat.meansDT <- DT::datatable(dat.means)
DT::saveWidget(dat.meansDT, 'dat.means.html')
cat('<iframe src="dat.means.html"style="border: none; width: 1200px; height: 1000px"></iframe>')
```

***

### Example: Cleaning Data
```{r dataTableExample3,eval=FALSE, cache=TRUE,cache.lazy = FALSE,warning=FALSE}
dim(dat.means)
# Set keys
setkey(dat.means, state)
dat.means <- dat.means[!c("AA","AE","PR","VI","ON","AP","GU")]
dat.means <- dat.means[nzchar(state)]

```

***

### Example: Making Interactive
```{r dataTableExample5,eval=FALSE, cache = FALSE,cache.lazy = FALSE,warning=FALSE, comment = NA,results='asis',error=FALSE}
library(rMaps)
library(htmlwidgets)
library(rcstatebin)
map <- statebin(as.data.frame(dat.means), payments ~ state,
                heading = "<b>Average Payment from Pharmaceutical Companies to Doctors By State</b>",
                footer = "<small>Data comes from openpaymentsdata.cms.gov</small>",
                type = "hex")

saveWidget(map,"meanstates.html", selfcontained = FALSE)


```

***

### Example: End Result
```{r,results='asis',comment=NA}
cat('<iframe src="./mean_states.html" width=100%, height=600></iframe>')
```

---  .class &vertical 

### rCharts: D3 in R

> - Create and customize interactive visualizations
> - Gives access to multiple D3 libraries
> - Easy to make and share

***

### A Lot from a Little

```{r tobaccoMapFake,cache=FALSE, results = 'asis', comment = NA,eval=FALSE}
library(rCharts)
library(knitr)
#read in csv
tobacco <- fread(
  "/home/steve/Documents/prettyGraphs/InteractiveGraphTutorial/tobacco.csv",header=TRUE)
# Source:
# https://research.stlouisfed.org/fred2/series/M0188AUSM149NNBR

# Convert characters to dates in seconds
  tobacco[, dates :=  as.numeric(
  as.POSIXct(tobacco$dates, origin="1929-06-14")) * 1000]

```

<!---
This actually runs
-->
```{r tobaccoMapReal,cache=FALSE, results = 'asis', comment = NA,echo=FALSE}
library(rCharts)
library(knitr)
#read in csv
tobacco <- fread(
  "/home/steve/Documents/prettyGraphs/InteractiveGraphTutorial/tobacco.csv",header=TRUE)
# Source:
# https://research.stlouisfed.org/fred2/series/M0188AUSM149NNBR

# Convert characters to dates in seconds
invisible(
  tobacco[, dates :=  as.numeric(
  as.POSIXct(tobacco$dates, origin="1929-06-14")) * 1000]
  )
```

```{r tobaccoChart2, cache=FALSE,warning=FALSE,echo=FALSE,results='asis'}
stargazer::stargazer(as.data.frame(tobacco$Tobacco_Consumption),type="html",flip = TRUE)
```
***

### Making FRED plots
```{r tobaccoChart,cache=FALSE}
# use rCharts to access Highcharts d3 library
h1 <- hPlot( 
  Tobacco_Consumption ~ dates,
  data = as.data.frame(tobacco),
  type = "line",
  title = "Tobacco Consumption, Manufactured Tobacco Supplies for United States"
           )
# Set type for x/y axis
h1$xAxis(type = "datetime")
h1$yAxis(title = list(text = "Tobacco Consumption in Millions lb."))
# Allow zoom
h1$chart(zoomType = "x")
h1$tooltip(formatter = # Kimm - how to make custom tooltip
              "#! function() { return 'Date: '     + 
                              Highcharts.dateFormat('%b/%e/%Y',
                                                    new Date(this.x)) +
                              '<br>' +'Consumption: ' + this.point.y; } !#")

```

`$object` comes from the [highchart's API](http://api.highcharts.com/highcharts) 

***

### Making FRED plots

```{r tobaccoMap2,cache=FALSE, results = 'asis', comment = NA,echo=FALSE}
h1$show('inline', include_assets = TRUE, standalone = TRUE)
```

***

### Making Bubble Chart

```{r, bubble, cache=FALSE,comment = NA,warning=FALSE}
a <- hPlot(Pulse ~ Height, data = MASS::survey, type = "bubble",
           title = "Zoom demo", subtitle = "bubble chart",
           size = "Age", group = "Exer")

a$chart(zoomType = "xy")
a$exporting(enabled = T)
```

***

### Making Bubble Chart

```{r,bubble2,cache=FALSE,results='asis',comment=NA,echo=FALSE}
a$show('inline',include_assets=TRUE,standalone= TRUE)
```


```{r,echo=FALSE}
hook1 <- function(x){ gsub("```\n*```r*\n*", "", x) }
hook2 <- function(x){ gsub("```\n+```\n", "", x) }
knit_hooks$set(document = hook2)
```

***

## Example App

<iframe src="https://johnricco.shinyapps.io/metro_walksheds/" style="border: none; width: 1200px; height: 900px"></iframe>


### NLP: Graph for topic model
[Link if iframe is broken](http://bl.ocks.org/Stevo15025/raw/d1f5a9ccbb36784ac9e2/#topic=0&lambda=1&term=now)

<iframe src="http://bl.ocks.org/Stevo15025/raw/d1f5a9ccbb36784ac9e2/#topic=0&lambda=1&term=now" style="border: none; width: 1200px; height: 1000px"></iframe>

---

## Resouces for Learning R

> - [Data Mining with R](http://www.dainf.ct.utfpr.edu.br/~kaestner/Mineracao/RDataMining/Data%20Mining%20with%20R-Kumar.pdf)
> - [Stackoverflow](http://stackoverflow.com/questions/tagged/r?sort=frequent&pageSize=50)
> - [NLP](http://onepager.togaware.com/TextMiningO.pdf)
> - [Datacamp](https://www.datacamp.com/courses/free-introduction-to-r)
> - [R-bloggers](http://www.r-bloggers.com/)


